---
title: "r_homework_2"
author: "Collin"
date: "4/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load Packages

```{r package}

library(tidyverse) 
library(rmarkdown) 
library(readxl)
library(janitor) 


```

##Load Data

```{r load data}

ac_data_raw <- read_xlsx("OCWR_AC_2019_Data.xlsx")

ac_data <- ac_data_raw %>% 
  clean_names()

ac_data$polygon_id[which(ac_data$polygon_id=="5M Buffer")] <- "5M BUFFER"

view(ac_data)

#names(ac_data)

oak_data_raw <- read_csv("Weir_Oak_Restoration_Data_winter19_2.csv")

oak_data <- oak_data_raw %>% 
  clean_names()

#view(oak_data)

#names(oak_data)

```

##Refer to Specific columns

```{r columns}

#display height column two ways

oak_data$height_cm

oak_data[4]

#display location_uml through sampling group

oak_data[6:8]

#Display columns 2, 5, 6, 7

oak_data %>% 
  select(2, 5, 6, 7)

#Display how to refer to every column except for 3 and 8

oak_data %>% 
  select(-3, -8)


```

##Change names of columns and observations

```{r change names}

#create new dataframe - oak_data2

oak_data2 <- oak_data

#view(oak_data2)

oak_data2$sampling_group[which(oak_data2$sampling_group=="L_C")] <- "lower_closed"

oak_data2$sampling_group[which(oak_data2$sampling_group=="M_C")] <- "middle_closed"

oak_data2$sampling_group[which(oak_data2$sampling_group=="U_C")] <- "upper_closed"

oak_data2$sampling_group[which(oak_data2$sampling_group=="L_O")] <- "lower_open"

oak_data2$sampling_group[which(oak_data2$sampling_group=="M_O")] <- "middle_open"

oak_data2$sampling_group[which(oak_data2$sampling_group=="U_O")] <- "upper_open"

colnames(oak_data2)[7] <- "supplementary"

#view(oak_data2)

## GOOGLE THIS LATER

#oak_data2$sampling_group[which(oak_data2$sampling_group %in% c("L_C", "M_C", "U_C", "L_O", "M_O", "U_O"))] <- c("lower_closed", "middle_closed", "upper_closed", "lower_open", "middle_open", "upper_open")
  



```


##Create new columns

```{r new_columns}


ac_data2 <- ac_data
ac_data2$pin_number_2 <- ac_data2$pin_number/2

#view(ac_data2)

ac_data3 <- ac_data2 %>% 
  mutate(new_column=native_non_native)

ac_data3$new_column[which(ac_data3$new_column=="Native")] <- "1"

ac_data3$new_column[which(ac_data3$new_column=="Non-Native")] <- "0"

ac_data3$new_column[which(ac_data3$new_column=="No_Native")] <- "."

ac_data3$pin_number_2_rounded <- round(ac_data3$pin_number_2)

#view(ac_data3)

```

##Create tables

```{r tables}

diversity1 <- ac_data %>% 
  filter(data_type ==	"T.PI") %>% 
  filter(native_non_native %in% c("Native", "Non-Native")) %>% 
  select(polygon_id, transect, native_non_native, species_code) %>% 
  group_by(polygon_id, transect, native_non_native) %>% 
  count(species_code)

#view(diversity1)

functional1 <- ac_data %>% 
  filter(native_non_native == "Native") %>% 
  filter(data_type ==	"T.PI") %>% 
  select(polygon_id, transect, functional_group) %>% 
  group_by(polygon_id, transect) %>% 
  count(functional_group)

#view(functional1)

```

##Practice with ggplots

#questions: How to group functional group labels by layer

```{r layer data}
layer_plot_data <- ac_data %>% 
  filter(native_non_native %in% c("Native", "Non-Native")) %>%
  select(layer, functional_group, species_code) %>% 
  group_by(layer, functional_group) %>% 
  count(species_code)

layer_plot_data$layer[which(layer_plot_data$layer==".")] <- "Non-Native"

layer_plot_data$functional_group[which(layer_plot_data$functional_group=="Forb")] <- "Non-Native Forb"

layer_plot_data$functional_group[which(layer_plot_data$functional_group=="Grass")] <- "Non-Native Grass"

layer_plot_data$functional_group[which(layer_plot_data$functional_group=="Shrub")] <- "Non-Native Shrub"

layer_plot_data[55, 1] = "Non-Native"

#view(layer_plot_data)

```

```{r sidebyside_plot}

View(layer_plot_data)

layer_plot_side <- ggplot(layer_plot_data, aes(fill = functional_group, y = n, x = layer)) + 
  geom_bar(position="dodge", stat="identity") + 
  theme_classic() + 
  ggtitle(label = "Functional Group Abundance by Layer")+
  xlab ("Layer")+
  ylab("Total Native Plant Count") +
  labs(fill = "Functional Group") +
  scale_x_discrete(expand=c(0.3,0))+
  scale_y_continuous(expand=c(0,0))+
  theme(axis.title.y = element_text(size = 10), plot.title = element_text(hjust = 0.5, face = "bold", size = 16)) + 
  scale_fill_manual(breaks=c("Annual forb\n(late flowering)", "Annual forb\n(other)", "Forb", "Grass", "Large Shrub", "Perennial Forb\n(late flowering)", "Perennial Forb\n(other)", "Perennial Grass\n(other)", "Shrub", "Shrub\n(late flowering)", "Shrub\n(other)", "Tree"), values = c("palegreen", "palegreen4", "turquoise", "firebrick1", "indianred2", "indianred3", "yellowgreen", "limegreen", "lawngreen", "darkorange4", "peru", "steelblue4"))


iris$Species <- factor(iris$Species, levels = c("virginica", "versicolor", "setosa"))

 scale_fill_discrete(breaks=c("trt1","ctrl","trt2"))  

layer_plot_side

```

##QUESTION - how do you change the order of the legend? There is one issue with the "herb" layer...one non-native point is labeled as herb not Non-Native - I manually fixed it but is there a better way?

```{r stacked_plot}

layer_plot_stack <- ggplot(layer_plot_data, aes(fill = functional_group, y = n, x = layer)) + 
  geom_bar(position="stack", stat="identity") + 
  theme_classic() + 
  ggtitle(label = "Functional Group Abundance by Layer")+
  xlab ("Layer")+
  ylab("Total Native Plant Count") +
  labs(fill = "Functional Group") +
  scale_x_discrete(expand=c(0.3,0))+
  scale_y_continuous(expand=c(0,0))+
  theme(axis.title.y = element_text(size = 10), plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = c("palegreen", "palegreen4", "turquoise", "firebrick1", "indianred2", "indianred3", "yellowgreen", "limegreen", "lawngreen", "darkorange4", "peru", "steelblue4"))

layer_plot_stack



```
```{r stacked_plot_normalized}

layer_plot_stacked_normalized <- ggplot(layer_plot_data, aes(fill = functional_group, y = n, x = layer)) + 
  geom_bar(position="fill", stat="identity") + 
  theme_classic() + 
  ggtitle(label = "Functional Group Abundance by Layer")+
  xlab ("Layer")+
  ylab("Functional Group %") +
  labs(fill = "Functional Group") +
  scale_x_discrete(expand=c(0.3,0))+
  scale_y_continuous(expand=c(0,0))+
  theme(axis.title.y = element_text(size = 10), plot.title = element_text(hjust = 0.5))+
  scale_fill_manual(values = c("palegreen", "palegreen4", "turquoise", "firebrick1", "indianred2", "indianred3", "yellowgreen", "limegreen", "lawngreen", "darkorange4", "peru", "steelblue4"))

layer_plot_stacked_normalized

```
```{r non_native}

ac_data_nn <- ac_data %>% 
  filter(native_non_native == "Non-Native") %>% 
  filter(data_type == "T.PI") %>% 
  count(species_code) %>% 
  filter(n >= 10) %>% 
  arrange(desc(n))

#view(ac_data_nn)


non_native_pie <- ggplot(ac_data_nn, aes(fill = species_code, y = n, x = "")) + 
  geom_bar(width=1, stat="identity") + 
  theme_void() + 
  coord_polar("y", start=0) + 
  geom_text(aes(label= n),position = position_stack(vjust = 0.5), color = "black", size=2.5) +
  scale_fill_brewer(palette="Set1") +
  ggtitle(label = "Non-Native Plants", subtitle= "Count of plants with 10 or more observations")+
  labs(fill = "Species Code") +
  theme(plot.title = element_text(hjust = 0.5, face="bold"), plot.subtitle=element_text(size=8, face="italic", hjust = 0.5), panel.grid = element_blank(), axis.title = element_blank(), axis.title.y = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.background = element_blank())
  
non_native_pie

  
  
  
```


## RK fixes

## Re-ordering legend items
```{r}

#Re-ordering legend items

layer_plot_side2 <- ggplot(layer_plot_data, aes(fill = functional_group, y = n, x = layer)) + 
  geom_bar(position="dodge", stat="identity") + 
  theme_classic() + 
  ggtitle(label = "Functional Group Abundance by Layer")+
  xlab ("Layer")+
  ylab("Total Native Plant Count") +
  labs(fill = "Functional Group") +
  scale_x_discrete(expand=c(0.3,0))+
  scale_y_continuous(expand=c(0,0))+
  theme(axis.title.y = element_text(size = 10), plot.title = element_text(hjust = 0.5, face = "bold", size = 16)) + 
  scale_fill_manual(values = c("palegreen", "palegreen4", "yellowgreen", "limegreen", "lawngreen", "turquoise", "steelblue4", "firebrick1", "indianred2", "indianred3", "darkorange4", "peru"), breaks=c("Annual Forb - Late Flowering","Annual Forb - Other","Perennial Forb - Late Flowering", "Perennial Forb - Other", "Perennial Grass - Other", "Large Shrub", "Tree", "Non-Native Forb", "Non-Native Grass", "Non-Native Shrub", "Shrub - Late Flowering", "Shrub - Other")) # Define breaks in the order you wish them to appear. Make sure to spell them perfectly. NOTE you must then reorder your colors accordingly

layer_plot_side2

```

## Fixing data errors
```{r fix}

#load data
ac_data_raw2 <- read_xlsx("OCWR_AC_2019_Data_example.xlsx") 
ac_data2 <- clean_names(ac_data_raw2)

View(ac_data_raw2)

#Use ifelse statement to create new column with fixed data
ac_data3 <- ac_data2 %>% 
  mutate(new_column = ifelse(native_non_native == "Non-Native" & layer == "herb",".",layer))
         
View(ac_data3)

```

























